Assortative mating MR summary

### 2021-09-15 analysis/AM_MR_summary.Rmd
add AM MR summary

tar_load(traits_final)
tar_load(household_MR_binned_het)
tar_load(household_MR_summary_AM)
tar_load(traits_corr2_filled)
tar_load(num_tests_by_PCs)
tar_load(num_tests_by_PCs_AM_sig)

@@

- For certain analyses it is worth only looking at MR results between couples where the exposure trait is the same as the outcome trait. 
- These results we have termed, the assortative mating MR results.

@@

- The number of analyses corresponds to the number of traits = `r dim(traits_final)[1]`. 
- The significant results after adjusting for multiple hypothesis testing are shown below. 
- `IVW_meta_beta` and `IVW_meta_pval` corresponds to the beta and p-value of the meta-analyzed MR across sexes, respectively.  
- After adjusting for multiple tests (p < `0.05 / `r num_tests_by_PCs``), the number of significant AM MRs was `r length(AM_sig_traits)` (see below on details for how number of effective tests was obtained).

@@

**A note on multiple testing adjustment:**
- Adjustment was based on PCs of correlation matrix.  
- Calculate correlation matrix of all `r dim(traits_final)[1]` traits tested. 
- Perform PCs on the resulting corr matrix. 
- Number of tests = # of PCs needed until variance explained is > 99.5%. 
- Resulting # of effective tests: `r num_tests_by_PCs`.

@@

- For each AM MR ($X_{p} \sim X_i$), analyses were run in each sex separately. 
- Of the `r length(AM_sig_traits)` significant results shown above, we tested to see if there was any difference in AM MR estimates between sexes. 
- A similar adjustment for multiple hypothesis testing was applied as above, resulting in the number of effective tests as `r num_tests_by_PCs_AM_sig`. 
- The table below displays the results that significantly differ among sexes (p < `0.05/`r num_tests_by_PCs_AM_sig``).

@@

- Only `r dim(AM_result_sig.DT)[1]/2` was significant after adjustment. 
- The table below shows the results significant at `p < 0.05`.

@@

@@

- For each AM MR ($X_{p} \sim X_i$), MR analyses were run in the full sample as well as in 5 roughly equal sized bins according to time couple had been together (`time_together`) estimated using [time at household variable](https://biobank.ctsu.ox.ac.uk/showcase/field.cgi?id=699), and median age (`age`). 
- Of the `r length(AM_sig_traits)` significant results shown above, we tested to see if there was any significant difference in AM MR estimates amongst the two grouping variables. 
- Specifically, the outcome data set was split into 5 bins, and the SNP-outcome effect was estimated in each bin separately. 
- These outcome-SNP effects were then used to generate bin-specific MR estimates, using the same SNP-exposure effects from Neale. 
- To estimate the difference among bins, we used two approaches: 
  1. The Cochran's Q test to test for heterogeneity in meta-analyses.
  2. Tested the significance of the slope of linear model of median bin (either `age` or `time together`) versus the bin-specific MR estimate.  
- The two results are shown below. Note that each statistic was computed in each sex separately and meta-analyzed.

@@

The number of significant q-statistics significant after multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig` = `r 0.05/num_tests_by_PCs_AM_sig```) in each group was:
  - Binned by median *age* of couples, number of significant results within meta = `r length(which(q_het_result_age$Q_pval_meta < 0.05/num_tests_by_PCs_AM_sig))/2` and sex-specific = `r length(which(q_het_result_age$Q_pval < 0.05/num_tests_by_PCs_AM_sig))/2`.
  - Binned by *time* couple had been together, number of significant results within meta = `r length(which(q_het_result_tt$Q_pval_meta < 0.05/num_tests_by_PCs_AM_sig))/2` and sex-specific = `r length(which(q_het_result_tt$Q_pval < 0.05/num_tests_by_PCs_AM_sig))/2`.

All results are displayed below (sort by Q_pval to view significant results):

@@

Slopes were calculated by estimating the beta-coefficient of a linear regression between MR estimate within each bin (outcome) and median bin (either age or time at same household). Linear models were run both *unweighted* and *weighted* for the by the inverse of the SE of the MR estimate.

The number of significant slopes ($MR_{bin} \sim median_{bin}$) significant after multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig` = `r 0.05/num_tests_by_PCs_AM_sig```) in each group was:
  - Binned by median *age* of couples, number of significant results within meta = `r length(which(slope_result_age$bin_slope_meta_pval < 0.05/num_tests_by_PCs_AM_sig))/2`, weighted by inverse of the SE = `r length(which(slope_result_age$bin_slope_meta_pval_wt < 0.05/num_tests_by_PCs_AM_sig))/2`, different between sexes = `r length(which(slope_result_age$bin_slope_sex_het_pval < 0.05/num_tests_by_PCs_AM_sig))/2`
  - Binned by *time* couple had been together, number of significant results within meta = `r length(which(slope_result_tt$bin_slope_meta_pval < 0.05/num_tests_by_PCs_AM_sig))/2` and weighted by inverse of the SE = `r length(which(slope_result_tt$bin_slope_meta_pval_wt < 0.05/num_tests_by_PCs_AM_sig))/2`, different between sexes = `r length(which(slope_result_tt$bin_slope_sex_het_pval < 0.05/num_tests_by_PCs_AM_sig))/2`
All results are displayed below (sort by Q_pval to view significant results):

All results are displayed below (sort by various p-values to view significant results):

@@

>>>>

tar_load(traits_final)
tar_load(household_MR_binned_het)
tar_load(household_MR_summary_AM)
tar_load(traits_corr2_filled)
tar_load(num_tests_by_PCs)
tar_load(num_tests_by_PCs_AM_sig)
tar_load(exposures_to_run)
tar_load(household_MR_AM_FvsM_fig)

@@

For certain analyses it is worth only looking at MR results between couples where the exposure trait is the same as the outcome trait. We've called these analyses *assortative mating MR (AM MR)*.

@@

- The number of analyses corresponds to the number of traits (`r dim(traits_final)[1]`). 
- The significant results after adjusting for multiple hypothesis testing are shown below. 
- `IVW_meta_beta` and `IVW_meta_pval` corresponds to the beta and p-value of the meta-analyzed MR across sexes, respectively (i.e. MR estimates were computed in each sex-seperately using sex-specific SNP-exposure and SNP-outcome results and then meta-analyzed).  
- After adjusting for multiple tests (p < `0.05/`r num_tests_by_PCs``), identified `r length(AM_sig_traits)` significant AM MR relationships (see below on details for how number of effective tests was obtained).

@@

**A note on multiple testing adjustment:**

  - Adjustment was based on PCs of correlation matrix.  
  - Calculate correlation matrix of all `r dim(traits_final)[1]` traits tested. 
  - Perform PCs on the resulting corr matrix. 
  - Number of tests = # of PCs needed until variance explained is > 99.5%. 
  - Resulting # of effective tests: `r num_tests_by_PCs`.

@@

- For each AM MR ($X_{p} \sim X_i$), analyses were run in each sex separately. 
- Of the `r length(AM_sig_traits)` significant results shown above, we tested to see if there was any difference in AM MR estimates between sexes. 
- A similar adjustment for multiple hypothesis testing was applied as above, resulting in `r num_tests_by_PCs_AM_sig` number of effective tests. 
- The table below displays the results that significantly differ among sexes (p < `0.05/`r num_tests_by_PCs_AM_sig``).

@@

Only `r dim(AM_result_sig.DT)[1]/2` was significant after adjustment. The table below shows the results significant at `p < 0.05`.

@@

In general, the male to female estimate is slightly larger than the female to male estimate, as shown in the figure below.

```{r sex_het_fig, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
household_MR_AM_FvsM_fig
```

@@

- For each AM MR ($X_{p} \sim X_i$), analyses were run in the full sample as well as in 5 roughly equal sized bins according to time couple had been together (`time_together_even_bins`) estimated using [time at household variable](https://biobank.ctsu.ox.ac.uk/showcase/field.cgi?id=699), and median age (`age_even_bins`). 
- Of the `r length(AM_sig_traits)` significant results shown above, we tested to see if there was any significant difference in AM MR estimates amongst the two grouping variables. 
- Specifically, the outcome data set was split into 5 bins, and the SNP-outcome effect was estimated in each bin separately (and each sex seperately). 
- These outcome-SNP effects were then used to generate bin-specific MR estimates, using the same SNP-exposure effects from Neale. 
- To estimate the difference among bins, we used two approaches: 
    1. The Cochran's Q test to test for heterogeneity in meta-analyses.
    2. Tested the significance of the slope of linear model of median bin (either `age` or `time together`) versus the bin-specific MR estimate.  
- The two results are shown below.

@@

The number of significant q-statistics significant after multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig` = `r 0.05/num_tests_by_PCs_AM_sig``) in each group was:

  - Binned by median *age* of couples, number of significant results within meta = `r length(which(q_het_result_age$Q_pval_meta < 0.05/num_tests_by_PCs_AM_sig))/2` and sex-specific = `r length(which(q_het_result_age$Q_pval < 0.05/num_tests_by_PCs_AM_sig))`.
  - Binned by *time* couple had been together, number of significant results within meta = `r length(which(q_het_result_tt$Q_pval_meta < 0.05/num_tests_by_PCs_AM_sig))/2` and sex-specific = `r length(which(q_het_result_tt$Q_pval < 0.05/num_tests_by_PCs_AM_sig))`.

All results are displayed below (sort by Q_pval to view significant results).

@@

Slopes were calculated by estimating the beta-coefficient of a linear regression between MR estimate within each bin (dependent variable) and median bin (independent variable, either age or time at same household). Linear models were run both *unweighted* and *weighted* for the by the inverse of the SE of the MR estimate.

The number of significant trends ($\beta$ estimates from the model: $\alpha_{bin} \sim median_{bin}$) significant after multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig` = `r 0.05/num_tests_by_PCs_AM_sig``) in each group was:

  - Binned by median *age* of couples, number of significant results after adjusting for number of tests:
      - $\beta$ unweighted, meta-anlayzed across sexes: `r length(which(slope_result_age$bin_slope_meta_pval < 0.05/num_tests_by_PCs_AM_sig))/2`
      - $\beta$ weighted by inverse of the SE, meta-anlayzed across sexes: `r length(which(slope_result_age$bin_slope_meta_pval_wt < 0.05/num_tests_by_PCs_AM_sig))/2`
      - $\beta$ significantly different between sexes: `r length(which(slope_result_age$bin_slope_sex_het_pval < 0.05/num_tests_by_PCs_AM_sig))/2`
  - Binned by *time* couple had been together, number of significant results after adjusting for number of tests:
      - $\beta$ unweighted, meta-anlayzed across sexes: `r length(which(slope_result_tt$bin_slope_meta_pval < 0.05/num_tests_by_PCs_AM_sig))/2`
      - $\beta$ weighted by inverse of the SE, meta-anlayzed across sexes: `r length(which(slope_result_tt$bin_slope_meta_pval_wt < 0.05/num_tests_by_PCs_AM_sig))/2`
      - $\beta$ significantly different between sexes: `r length(which(slope_result_tt$bin_slope_sex_het_pval < 0.05/num_tests_by_PCs_AM_sig))/2`.

All results are displayed below (sort by various p-values to view significant results).

```{r bin_slope, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

@@

### Figures

A few figures of interest are shown below.

```{r binned_figures, include = T, echo = FALSE}

target1 <- which(exposures_to_run$Neale_pheno_ID=="1080")
figs1 <- tar_read(household_MR_binned_AM_figs, branches = target1)[[1]]

figs1[[5]]
figs1[[7]]

target2 <- which(exposures_to_run$Neale_pheno_ID=="1239")
figs2 <- tar_read(household_MR_binned_AM_figs, branches = target2)[[1]]

figs2[[5]]
figs2[[7]]

target3 <- which(exposures_to_run$Neale_pheno_ID=="20153_irnt")
figs3 <- tar_read(household_MR_binned_AM_figs, branches = target3)[[1]]
figs3[[5]]
figs3[[9]]

target4 <- which(exposures_to_run$Neale_pheno_ID=="1050")
figs4 <- tar_read(household_MR_binned_AM_figs, branches = target4)[[1]]
figs4[[4]]
figs4[[8]]
```
